---
title: "Projects"
subtitle: ""
author: ""
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    selfcontained: true
    # mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    # mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_SVG"
    # css: "kunoichi"
    css: xaringan-themer.css
    seal: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
# bibliography: ../../paper-rmd/assets/voter-id-bib.bib
# biblio-style: ../../paper-rmd/assets/apsr-leeper.bst
---

class: middle, center


```{r setup-rmd, include = FALSE, cache = FALSE}
source(here::here("code", "assets", "slide-helpers.R"))
```

```{r inf, include = FALSE, cache = FALSE, eval = FALSE}
xaringan::inf_mr(here::here("lessons", "01-intro", "01-intro.rmd"))
# servr::daemon_stop(1)
```


.pull-left[
# **PS 811: Statistical Computing**

<br>
Michael DeCrescenzo
<br>
University of Wisconsin–Madison

]

.pull-right[
```{r, include = TRUE, fig.width = 4, fig.height = 3, out.width = "100%"}
tibble(
  x = seq(0, 1, .01)
) %>%
  crossing(b = seq(1, 10, length.out = 5)) %>%
  mutate(dbeta = dbeta(x, 3, b)) %>%
  ggplot() +
  aes(
    x = x, 
    y = dbeta, 
    color = as.factor(b), 
    fill = as.factor(b)
  ) +
  geom_line() +
  geom_ribbon(
    aes(ymin = 0, ymax = dbeta),
    color = NA,
    alpha = 0.2
  ) +
  viridis::scale_fill_viridis(discrete = TRUE, end = 0.9) +
  viridis::scale_color_viridis(discrete = TRUE, end = 0.9) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank()
  ) 
```

]

## Projects, R Markdown, basic R

`r format(Sys.time(), '%B %d, %Y')`



---

# [github.com/mikedecr/example-project](https://github.com/mikedecr/example-project)

--

Download as `.zip` (save somewhere that ISN'T your PS-811 folder)

or, in your command line terminal (e.g. in RStudio)

```{bash, eval = FALSE, include = TRUE, echo = TRUE}
# navigate to a folder of your choosing, and then
# clone (download/copy) the repository

pwd              # (P)rint (w)orking (d)irectory (aka current folder)
ls               # (L)i(s)t folder contents
cd ~/desktop     # (C)hange (D)irectory to place of your choosing
                 # ~ means "top of user profile"
                 # [this particular destination may not work for you!]

# copy this url from the Github page
git clone https://github.com/mikedecr/example-project.git
```


---

# Open `project-demo.Rproj`

--

.pull-left[

The "Panes"

- **Console**: R's living brain. Where commands are executed/printed.

- **Source**: Write code, save files. Execute w/ a "keybind"

- **Environment:** What data objects do I have?

- **Files/Plots/Help**: Useful misc. things

]

--

.pull-right[

Why `.Rproj?`

- Top right? Project name!
- Files? The project folder!

```{r, echo = TRUE, include = TRUE, eval = FALSE}
# type in console:
getwd()

# prints current location 
# (hint: the project folder!)
```

]


---

# Projects in Action

In the "Files" pane, open `00_build.R`

Install the packages (one at a time) 

- This also updates package versions

- Packages? Contain extended R capabilities (functions)


While these install...



---

# Why Projects?

.pull-left[

**Organization**: everything in its place

**Replication**: coherent structure means _robust to development_

]


.pull-right[

**Portability**: self-contained, can be moved/shared

**Transparency**: sync to Github, view changes/contributors, rewind

]


---

class: middle, center

# Continue building project


---

# Project-oriented practices

.pull-left[

**R should be a blank slate**

- New R "session" when (re-)opening a project
- Never _save_ nor _restore_ session `.Rdata` (change settings)
- Save the _source code_, not the results! (REPRODUCIBILITY)
- Leave a paper trail: NEVER modify data by hand, NEVER save over raw data, ONLY use code

]

--

.pull-right[

**Open RStudio from `.Rproj` file**


**"Practice safe paths"**

- What is a path?
- To-down view: project "root"
- Subfolders are for going _in_, never _out_
- `file.path("data", "clean")`
- `here("data", "clean")`


Use [**good names**](https://speakerdeck.com/jennybc/how-to-name-files) (link)

]


---

class: center, middle

# Start a new RStudio project for this course

Already have a PS-811 folder? Too bad. Start a new one.


---

# Two types of files

.pull-left[
**R file**

- file extension `.R`

- contains ONLY R code

- send commands directly to console
]

.pull-right[
**R Markdown**

- file extension `.Rmd`, combines R code with writing ("Markdown" syntax)

- When editing, can still write/execute R code within "code chunks"

- Intended to be converted to your desired output file type (HTML, PDF)

- Markdown: plain text symbols that convey formatting of the output (e.g. `_italics_` or `**bold**`)

]

---

# Rendering ("knitting") an R Markdown document

When you "knit," this happens automatically:

1. Begin with `.Rmd` file, containing R and writing. 

2. R code is evaluated. Results incorporated into the rest of the Markdown (`.md`). 

3. Sent to Pandoc, a program that converts document formats.
    - HTML is "default" (no output specified)
    - if you choose PDF, this invokes $\mathrm{\LaTeX}$

---

class: middle, center

# New R file, new R Markdown file







