---
title: "Regression"
subtitle: ""
author: ""
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    selfcontained: true
    # mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    # mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_SVG"
    # css: "kunoichi"
    css: xaringan-themer.css
    seal: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
# bibliography: ../../paper-rmd/assets/voter-id-bib.bib
# biblio-style: ../../paper-rmd/assets/apsr-leeper.bst
---

class: middle, center


```{r setup-rmd, include = FALSE, cache = FALSE}
source(here::here("code", "assets", "slide-helpers.R"))
```

```{r inf, include = FALSE, cache = FALSE, eval = FALSE}
xaringan::inf_mr(here::here("slides", "04_regression", "04_regression.rmd"))
# servr::daemon_stop(1)
```


.pull-left[
# **PS 811: Statistical Computing**

<br>
Michael DeCrescenzo
<br>
University of Wisconsin–Madison

]

.pull-right[
```{r, include = TRUE, fig.width = 4, fig.height = 3, out.width = "100%"}
tibble(
  x = seq(0, 1, .01)
) %>%
  crossing(b = seq(1, 10, length.out = 5)) %>%
  mutate(dbeta = dbeta(x, 3, b)) %>%
  ggplot() +
  aes(
    x = x, 
    y = dbeta, 
    color = as.factor(b), 
    fill = as.factor(b)
  ) +
  geom_line() +
  geom_ribbon(
    aes(ymin = 0, ymax = dbeta),
    color = NA,
    alpha = 0.2
  ) +
  viridis::scale_fill_viridis(discrete = TRUE, end = 0.9) +
  viridis::scale_color_viridis(discrete = TRUE, end = 0.9) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank()
  ) 
```

]

## Regression and Model Output

February 28, 2020

---

class: middle

.pull-left[
# Goals

1. Estimating common regressions models

2. Understanding/visualizing model output

3. Generating "post-estimation" quantities
]



--

.pull-right[
# Translated

1. Estimating parameters

2. Visualizing parameters

3. _Functions of_ parameters

]

---

class: inverse, middle, center

## In your code-along file...


---

# Packages we have worked with

.pull-left[

```{r cache = FALSE, include = TRUE, results = 'hide', echo = TRUE}
library("here") 
library("tidyverse")
library("broom")
```

]

.pull-right[

**here**: manage file pathways safely

**tidyverse**: bottomless pit of tools for data manipulation/graphics

**broom**: turning model output into _data frames_

]

--

.center[
<br>

### (special appearances by other packages)
]


---

# New data on Canvas

```{r,  echo = TRUE, include = TRUE, results = 'hide'}
house <- 
  here("data", "house-elections.csv") %>%
  read_csv() %>% 
  print()
```

--

.pull-left[

Data on House members since 1900s

- NOMINATE score from -1 (most liberal) to +1 (most conservative)
- (new) Previous **House** vote outcome in their district
- (new) Previous **presidential** vote outcome in their district
]


--

.pull-right[
Vote outcomes measured as Republican share of the **"two-party vote"**

\begin{align}
  \text{Rep. share} &= \frac{\text{Rep. Votes}}{\text{Rep. Votes} + \text{Dem. Votes}} 
\end{align}
<!--  -->
]



---

# $y_{i} = \alpha + \beta x_{i} + \epsilon_{i}$

`lm()` for linear model

```{r include = TRUE, echo = TRUE, results = "hide"}
# lm(formula = y ~ x, data = data_name)
simple_reg <- lm(nominate_dim1 ~ rep_share_house, data = house)
```

--

```{r include = TRUE, echo = FALSE}
simple_reg
```


---

# Why `broom`?

.pull-left[

The old way: 

- `summary(model)`

- `confint(model)`

- `predict(model, interval = "confidence")`

- Works, but clunky. Must covert to other object classes to make useful

]

--


.pull-right[

The cool way: the `{broom}` package

- does all of the conversion for you

- results are **data frames**

- natural compatibility with `{tidyverse}`

]


---

# `broom::tidy()`

Summarizes the model at the **variable level** (coefficients, std. error, $p$-value)

```{r include = TRUE, echo = TRUE, results = "hide"}
tidy(simple_reg)
```

--

```{r include = TRUE, echo = FALSE}
tidy(simple_reg)
```


---


# `broom::glance()`

Summarizes the model at the **model level** ( $R^{2}$, $\sigma$, $F$, DFs)

```{r include = TRUE, echo = TRUE, results = "hide"}
glance(simple_reg)
```

--

```{r include = TRUE, echo = FALSE}
glance(simple_reg)
```



---

# `broom::augment()`

Summarizes the model at the **observation level** ( $x_{i}$, $y_{i}$, conditional mean, residuals)

```{r include = TRUE, echo = TRUE, results = "hide"}
augment(simple_reg)
```

--

```{r include = TRUE, echo = FALSE}
augment(simple_reg)
```


---

# Plot residuals

.pull-left[

```{r rvf, include = TRUE, echo = TRUE, results = "hide", fig.show = 'hide'}
augment(simple_reg) %>%
  ggplot() +
  aes(x = rep_share_house, 
      y = .resid) +
  geom_point()
```

]

.pull-right[

```{r rvf, include = TRUE, fig.width = 4, fig.height = 3, out.width = "100%"}
```

]



---

# Multiple regression

$Y = X\boldsymbol{\mathbf{\beta}} + \mathbf{E} \qquad$ or $\qquad y_{i} = \mathbf{x}_{i}^{\intercal}\boldsymbol{\mathbf{\beta}} + \epsilon_{i}$

```{r multi, include = TRUE, echo = TRUE, results = "hide"}
# also note PIPE INTO NON-FIRST ARGUMENT
# use `.` to stand-in for data
multi_mod <- house %>%
  lm(nominate_dim1 ~ rep_share_house + republican, data = .)

tidy(multi_mod)
```

--

```{r multi, include = TRUE}
```




---

# See notes for how-tos

.pull-left[

Model formula tricks:

- Suppress the constant/intercept

- convert categories to indicator variables

]

--

.pull-right[

Other types of models:

- general linear models (e.g. logistic model)

- heteroskedasticity-consistent (a.k.a. "robust") standard errors

]


---

# Coefficient plot. Estimate, tidy, plot

```{r}
filter <- dplyr::filter
```

```{r coef-data, include = TRUE, echo = TRUE}
bigger_tidy <- house %>%
  filter(election_year >= 2000) %>%
  lm(formula = nominate_dim1 ~ scale(rep_share_house) +
               scale(rep_share_pres) + republican +
               as.factor(election_year), 
     data = .) %>%
  tidy(conf.int = TRUE) %>%
  print() 
```

---


```{r coefs, include = TRUE, echo = TRUE, out.width = "100%", fig.width = 8, fig.height = 3}
ggplot(bigger_tidy) +
  aes(x = term, y = estimate) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0) +
  coord_flip()
```

---

# Why figures?

- easier to see than to read

- combine many models in same figure (stack data frames with `bind_rows`)

- easy to get a sense of scale

- less stargazing (magnitude ≠ significance)

- plot effect sizes rather than coefficients




---


.pull-left[
# Tabular output

- Save time/effort
- Prevent _transcription errors_
- Update automatically

```{r htmlreg, include = TRUE, echo = TRUE, results = "hide"}
texreg::htmlreg(
  multi_mod, 
  doctype = FALSE,
  caption = "TeXreg Demo"
)
```

See notes for more

]

.pull-right[

```{r htmlreg, include = TRUE, echo = FALSE, results = "asis"}
```
]


---


.pull-left[

# Predictions

- `augment()` takes `newdata = ` argument
- supply custom data to make predictions for
- make a data frame with `tibble` or `tribble`
- trick: use `crossing()`

]

--

.pull-right[

```{r, include = TRUE, echo = TRUE}
pred_data <- 
  crossing(
    rep_share_house = 0:100,
    republican = c(0, 1)
  ) %>%
  print()
```

]

---

# Pass new data to augment

```{r include = TRUE, echo = TRUE}
augment(multi_mod, newdata = pred_data)
```


---

# Make confidence intervals

```{r include = TRUE, echo = TRUE}
predictions <- 
  augment(multi_mod, newdata = pred_data) %>%
  mutate(
    conf.low = .fitted - (.se.fit * 1.96), # should use residual DFs
    conf.high = .fitted + (.se.fit * 1.96) # from glance() results
  ) %>%                                    # to get t statistic
  print()
```

---


# Plot predictions

.pull-left[

```{r plot-aug, include = TRUE, echo = TRUE, fig.show = "hide"} 
ggplot(predictions) +
  aes(x = rep_share_house, y = .fitted, 
      fill = as.factor(republican)) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    color = NA, alpha = 0.25
  ) +
  geom_line() +
  theme(legend.position = "none") +
  scale_fill_manual(
    values = c("0" = "dodgerblue", "1" = "tomato")
  )
```
]

.pull-right[
```{r plot-aug, include = TRUE, fig.width = 4, fig.height = 3, out.width = "100%"}
```
]

---

class: center

# Closing notes on regression

**There is no one path** through regression in R.

**What `broom` gives us** is tools that simplify our workflow

**`glance()` and `tidy()`** will last, but **you will outgrow `augment()`** if you need more complicated post-estimation quantities

**See the notes** for ideas on where to look next

**No substitute for knowing the statistics** even if packages help with the details





