---
title: |
  | **SOLUTIONS** 
  | Skill Task 1: Rmd, `here()`
author: "PS 811: Statistical Computing"
date: |
  | `r format(Sys.time(), "%B %d, %Y")`
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: pdflatex
    toc: false
    number_sections: false
geometry: margin=1.25in
fontsize: 12pt
subparagraph: yes
citecolor: black
urlcolor: blue
linkcolor: magenta
# latex packages follow:
# - Linux Libertine for text and math fonts 
# - Inconsolata for code
header-includes:
  \usepackage{libertine}
  \usepackage{libertinust1math}
  \usepackage[varqu, scaled = 0.95]{zi4}
---


<!-- Hi! I am a Markdown comment. -->

<!-- 
You can break a markdown comment
across
multiple lines
 -->


```{r chunks, include = FALSE}
# show code and R output in the same code block
knitr::opts_chunk$set(collapse = TRUE)
```



# Setup

In this class, we take advantage of RStudio's "Project" capabilities. We open an R project by opening the `.Rproj` file for a project, or opening the desired project from within RStudio. 

When you work within an R project, R's working directory should automatically be set to the project "root," a.k.a.\ the top of the project folder. You can check this using the `here` package, which gives you feedback about its behavior when you load it.

```{r load-here}
library("here")
```

If this does print your project directory, two things could have gone wrong: (1), you aren't working in your RStudio project, or (2) you have not given your project an `.Rproj` file.


# Basic R commands

I create a code chunk that saves two variables, `first` and `last`.

```{r save-name}
first <- "Michael"
last <- "DeCrescenzo"
```

I now use the `c()` function to combine my first and last name into one vector. The `c()` function is ubiquitous in R. Any time you need to supply a _series of values_ to a function argument, for example, you can create that series (a.k.a.\ vector) using `c()`.

```{r combine-name}
c(first, last)
```

# Import data

Now I import the `CAFE` data into R. I need to load the `tidyverse` package to get the `read_csv()` function.^[
  You should be aware that `read_csv()`, a function contained within `tidyverse`, is different from `read.csv()`, a function that R has by default. We will generally prefer reading functions whose names have _underscores_, so be mindful.
]

```{r load-tidy}
library("tidyverse")
```

When you load the `tidyverse` package, it will loudly tell you a few things. First, because `tidyverse` is technically a bundle of packages, it tells you that several packages have actually been loaded. They are:

- `readr`: data import/export functions. Technically this is where `read_csv()` lives! 
- `tibble`: improves the behaviors of data frames in R. [Learn more.](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html)
- `dplyr`: data manipulation functions. We've seen this already!
- `ggplot2`: graphics. We've also seen this!
- `tidyr`: a package for data reshaping. This package is an update to other packages that you may have seen before, `reshape` and `reshape2`. We will explore this soon!
- `stringr` and `forcats`: two packages that provide helpful functions for working with character strings and factor data. We will also explore this soon!
- `purrr`: tools for working with lists, list-columns, and nested data. If this sounds like "wtf" stuff, that's because it's advanced. But buckle up because we're going to dive in at the end of the semester.

There is also a message that tells you that functions have been "masked." What this means is that there are functions within `tidyverse` packages that have the same names as other R functions, and so the new functions are prioritized over the old functions. Never fear; you can always access a masked function by typing `package_name::function_name()`.

Now I read the data using `read_csv()`, which requires a path to the file that we want to import. We can always check that our dataset is where it's supposed to be by using R to look within our data folder. You can print folder contents (file names and folder names) using `list.files()`.^[
  If you find yourself in a situation where data import isn't working, you can use `list.files()` to check your assumptions. Sometimes you mistype little things, so you can even copy and paste the output to ensure that everything is spelled correctly.
]



```{r list-files}
# print the contents of the project folder
list.files(here())

# print the contents of the data folder
list.files(here("data"))
```

We use `here()` to build the path, so we call `here()` within `read_csv()`. The functions are evaluated inside out, so `here()` builds the path, and then `read_csv()` uses the path to import the data. Note that I also pipe into `print()`, which lets me look at the results even as I store the data in the object named `cafe`.

```{r read-data}
cafe <- 
  read_csv(here("data", "CAFE.csv")) %>%
  print()
```

Reading data with `read_csv()` (and other `tidyverse`-style `read_*()` functions) will also print a noisy message. It isn't an error. It's telling you how the data were interpreted by `read_csv()`. It's saying, "I made a variable called 'Senator' that is a character (string) variable..." and so on. These are inferences that the function is making about the data, but it's possible to modify this inferential behavior in advanced applications.

If your code looks like mine, and the document builds ("knits") to PDF, then you did great! 

